---
title: "Analysis plan"
author: "Lucija Batinović"
format: html
editor: source
---

## Model specifications

As the reporting conventions of single-case designs allow for raw data extraction, we will take advantage of this opportunity and conduct an individual-participant multilevel meta-analysis. Having access to raw data allows for greater flexibility in modeling variations between and within participants, which is of even greater importance in single-case designs, considering the low sample size and often very few data points. For that reason, we will conduct a Bayesian meta-analysis. 

There are multiple crucial assumptions to make when modeling time-series (i.e., repeated measures) designs: distribution of the outcome measure, time trends, autoregression, etc. There is little consensus on how to best approach dealing with outcome measures in single-case designs. Most commonly used effect size measures in SCD research belong to non-parametric effect measures, e.g., PND (percentage of non-overlap data), LRR (logistic response ratio) and more recently developed parametric alternatives, e.g., WS-SMD and BS-SMD (standardized mean differences; developed to be comparable to group research). These parametric effect size measures, although robust when assumptions cannot be certainly defined, generally assume non-existing time trends and autoregression, with a normal distribution of outcome measures (cite Pustejovsky). Shadish (cite the scd study) found that autocorrelation size is inconclusive, albeit non-negligible, particularly in multiple-baseline designs. Furthermore, assuming normal distribution of the outcome measures is a risky feat in a research area that heavily utilizes counts and percentages as measures, which would be more appropriately explained under the Poisson family of distributions. We will further break down each parameter that will go into the model, justifying the final decisions for each one.

#### Outcome
Level one of the multilevel model describes the relationship between the outcome - word or non-word reading, and the predictor - intervention. Informed by previous reviews with similar interest, we assume majority of the studies will employ non-standardized word-reading tests, with counts and percentages as measures, and fixed word lists. If we considered the trials to be independent, binomial or Poisson distribution would potentially be the most appropriate model for the outcome. However, as we assume the existence of autocorrelation and knowledge transfer, we cannot assume the trials are independent. To allow for greater flexibility and account for over-dispersion (i.e., variance being larger than the mean), the outcome will be modeled under the negative binomial distribution. 

This sets the beginning function as:

$$
Y_{ij} = \beta_0 + \beta_1*Phase + \epsilon
$$
$$
\epsilon \sim N(0,\sigma^2)
$$
Where: $Y_{ij}$  is the outcome for _i_ case in _j_ study. $\beta_0$  represents the intercept, which is the baseline phase, and $\beta_1$  represents the slope, which is the change in the outcome after the intervention is implemented. When the phase is 0, the outcome is the intercept, i.e., the baseline level. 

```{r}
#| echo = FALSE
library(tidyverse)
library(brms)
```

```{r data}
#| echo = FALSE
data <- read_csv("studies_meta.csv") %>% mutate(across(c(1:5), ~as_factor(.)))

```

```{r baes-model}
#| echo = FALSE


#baes <- brms::brm(formula = y ~ phase + session + ( 1 | phase/session) + #ar(time = session, p = 1), #fixed effect within case; autoregression order 1
#                       family = negbinomial(),# neg binomial
#                       data = data,
#                  seed = 2023) # 
#
#summary(baes)

```

```{r freaks-model}
#| echo = FALSE
#freaks <- brms::brm(formula = outcome ~ phase + session + ( 1 | phase/session) + ar(time = session, p = 1), #fixed effect within case; autoregression order 1
#                     family = gaussian(), # outcome distribution considered gaussian
#                     data = data) 

#summary(freaks)
```

